\pkg{casebase}: An Alternative Framework For Survival Analysis

### 2020-03-20 analysis/paper/paper.Rmd
Auto stash before merge of "write-popTime" and "origin/master"

1. The \code{casebase::popTime} function takes as input the original dataset along with the column names corresponding to the timescale, the event status and an exposure group of interest (optional). This will create an object of class \code{popTime} (or \code{popTimeExposure} if you specify a value for the exposure argument).  
2. The corresponding \code{plot} methods for the object created in Step 1 can be called to create the population time plot with several options for customizing the aesthetics.

@@

```{r plot-erspc-data, eval = eval_cs1}
data("ERSPC")
ERSPC$ScrArm <- factor(ERSPC$ScrArm, 
                       levels = c(0,1), 
                       labels = c("Control group", "Screening group"))
pt_object <- casebase::popTime(ERSPC, 
                               time = "Follow.Up.Time",
                               event = "DeadOfPrCa",
                               exposure = "ScrArm")
attr(pt_object, "exposure")
```

@@

We can also create exposure stratified plots by specifying the \code{exposure} argument in the \code{popTime} function:

```{r plot-stratified-erspc-data, eval = eval_cs1}
pt_object_strat <- casebase::popTime(ERSPC, 
                                     event = "DeadOfPrCa", 
                                     exposure = "ScrArm")
# plot(pt_object_strat)
```

@@

```{r, eval = F}
knitr::knit_exit()
```

>>>>

The splitting of these tasks gives the user the flexibility to create population-time plots with the graphics system of their choice using the object created in Step 1, while the object created in Step 2 allows one to leverage the \pkg{ggplot2} family of functions for further customization. 
1. The \code{casebase::popTime} function takes as input the original dataset along with the column names corresponding to the timescale, the event status and an exposure group of interest (optional). This will create an object of class \code{popTime} (or \code{popTimeExposure} if you specify a value for the exposure argument).  
2. The corresponding \code{plot} methods for the object created in Step 1 can be called to create the population time plot with several options for customizing the aesthetics.

@@

We first pass this dataset to the \code{popTime} function. If you do not specify the time and event arguments, this function will try to guess them using regular expressions. We then generate the plot by using the corresponding \code{plot} method:
```{r plot-erspc-data, eval = eval_cs1}
data("ERSPC")
ERSPC$ScrArm <- factor(ERSPC$ScrArm, 
                       levels = c(0,1), 
                       labels = c("Control group", "Screening group"))
pt_object <- casebase::popTime(ERSPC, 
                               time = "Follow.Up.Time",
                               event = "DeadOfPrCa",
                               exposure = "ScrArm")
attr(pt_object, "exposure")
```

@@

The randomization of the Finnish cohorts were carried out on January 1 of each of the 4 years 1996 to 1999. This, coupled with the uniform December 31 2006 censoring date, lead to large numbers of men with exactly 11, 10, 9 or 8 years of follow-up. Tracked backwards in time (i.e. from right to left), the population-time plot shows the recruitment pattern from its beginning in 1991, and in particular the Jan 1 entries in successive years. Tracked forwards in time (i.e. from left to right), the plot for the first three years shows attrition due entirely to death (mainly from other causes). Since the Swedish and Belgian centres were the last to close their recruitment - in December 2003 - the minimum potential follow-up is three years. Tracked further forwards in time (i.e. after year 3) the attrition is
We can also create exposure stratified plots by specifying the \code{exposure} argument in the \code{popTime} function:

```{r plot-stratified-erspc-data, eval = eval_cs1}
pt_object_strat <- casebase::popTime(ERSPC, 
                                     event = "DeadOfPrCa", 
                                     exposure = "ScrArm")
# class(pt_object_strat)
# pt_object
plot(pt_object_strat)
# dev.off()
# library(magrittr)
# pt <- plot(pt_object_strat)
# pt %>% class

# library(ggplot2)
# pt + theme_minimal()

```

@@

```{r, eval = T}
knitr::knit_exit()
```
